Gametic Selection, Meiotic Drive, Sex Ratio Bias, and Transitions Between Sex Determination Systems

### 2017-05-19 LaTeX/Turnover.tex
table 2 updated and appenidx reworded

\textcolor{red}{I now remove the zygotic sex ratio $\zeta$ from the mean fitnesses. The mean fitnesses will have to be adjusted in a corresponding way.}

\textcolor{red}{CHECK: DO THE HAPLOID MEAN FITNESSES HAVE TO BE IN THE DENOMINATOR? i think so, based on definitions in table S.2 (mmo)}

@@

\begin{threeparttable}[ht]
\centering
\smallskip
%\caption{Parameters determining invasion (equation \ref{eq:charpoly_neoY}) for neo-Y or neo-W alleles}
\caption{Parameters determining invasion of mutant neo-Y and neo-W alleles into an ancestrally XY system}
\begin{tabular}{l}
\hline\hline
   \noalign{\vskip 0.5ex}
   neo-Y ($k=0$) \\ [0.5ex] \hline \noalign{\vskip 1ex}
  $\lambda_{mA}=[1/(2(1- \zeta))]\{p_{X}^{\female} w_{A}^{\female} w_{A}^{\male} w_{AA}^{\male} + 2(1-p_{X}^{\female}) w_{a}^{\female} w_{A}^{\male} w_{Aa}^{\male} \alpha^\male (1 - R)\}/ \bar{w}^\male $\\ [0.5ex] \noalign{\vskip 0.5ex}
  $\lambda_{ma}=[1/(2(1- \zeta))]\{(1-p_{X}^{\female}) w_{a}^{\female} w_{a}^{\male} w_{aa}^{\male} + 2p_{X}^{\female} w_{A}^{\female} w_{a}^{\male} w_{Aa}^{\male}(1 - \alpha^\male) (1 - R)\}/ \bar{w}^\male $ \\ [0.5ex] \noalign{\vskip 0.5ex}
  $\rho_{mA}=R (1-p_{X}^{\female}) w_{a}^{\female} w_{A}^{\male} w_{Aa}^{\male} \alpha^{\male}/  (1-\zeta) \bar{w}^\male  $\\ [0.5ex] \noalign{\vskip 0.5ex}
  $\rho_{ma}=R p_{X}^{\female} w_{A}^{\female} w_{a}^{\male} w_{Aa}^{\male} (1 - \alpha^{\male})/ (1-\zeta) \bar{w}^\male $\\ [1ex] \hline 
  \noalign{\vskip 0.5ex}
  neo-W ($k=1$) \\ [0.5ex] \hline \noalign{\vskip 1ex}
  $\lambda_{mA}=[1/(2 \zeta)]\{\bar{p}^{\male} w_{A}^{\male} w_{A}^{\female} w_{AA}^{\female}+2(1-\bar{p}^{\male}) w_{a}^{\male} w_{A}^{\female} w_{Aa}^{\female}\alpha^\female(1-R)\}/ \bar{w}^\female $ \\ [0.5ex] \noalign{\vskip 0.5ex}
  $\lambda_{ma}=[1/(2 \zeta)]\{(1-\bar{p}^{\male}) w_{a}^{\male} w_{a}^{\female} w_{aa}^{\female}+2\bar{p}^{\male} w_{A}^{\male} w_{a}^{\female} w_{Aa}^{\female}(1-\alpha^\female)(1-R)\}/ \bar{w}^\female $ \\ [0.5ex] \noalign{\vskip 0.5ex}
  $\rho_{mA}=R (1-\bar{p}^{\male}) w_{a}^{\male} w_{A}^{\female} w_{Aa}^{\female} \alpha^{\female}/ \zeta \bar{w}^\female  $\\ [0.5ex] \noalign{\vskip 0.5ex}
  $\rho_{ma}=R\bar{p}^{\male} w_{A}^{\male} w_{a}^{\female} w_{Aa}^{\female} (1-\alpha^{\female})/ \zeta \bar{w}^\female $\\ [1ex] 
  \hline \hline 
   \end{tabular}
      \begin{tablenotes}
      \scriptsize
      \item $\bar{p}^{\male}=p_{Y}^{\male}q+p_{X}^{\male}(1-q)$ is the average frequency of the $A$ allele among X- and Y-bearing male gametes.
      \item $R$ is the probability of recombination between loci \textbf{A} and \textbf{M}.
      \item $\zeta$ is the zygotic sex ratio and $\bar{w}^\Hermaphrodite$ is the mean fitness of sex $\Hermaphrodite$, see Table \ref{tab:meanfitnesses}.
%      \item $\bar{w}_H^\female = p_X^\female w_A^\female + (1-p_X^\female) w_a^\female$ is the mean fitness of female gametes during gametic competition
%      \item $\bar{w}_H^\male = \bar{p}^{\male} w_A^\male + (1-\bar{p}^{\male}) w_a^\male$ is the mean fitness of male gametes during gametic competition
%       \item $\bar{w}^\female = \{ p_X^\female w_A^\female (1-q) p_X^\male w_A^\male w_{AA}^\female + (1 - p_X^\female) w_a^\female (1-q) p_X^\male w_A^\male w_{Aa}^\female + p_X^\female w_A^\female (1-q) (1 - p_X^\male) w_a^\male w_{Aa}^\female + (1-p_X^\female) w_a^\female (1-q) (1 - p_X^\male) w_a^\male w_{aa}^\female \} / \{ \bar{w}_H^\female \bar{w}_H^\male\}$ is the mean fitness of females during diploid competition
%       \item $\bar{w}^\male = \{ p_X^\female w_A^\female q p_Y^\male w_A^\male w_{AA}^\male + (1 - p_X^\female) w_a^\female q p_Y^\male w_A^\male w_{Aa}^\male + p_X^\female w_A^\female q (1 - p_Y^\male) w_a^\male w_{Aa}^\male + (1-p_X^\female) w_a^\female q (1 - p_Y^\male) w_a^\male w_{aa}^\male \} / \{ \bar{w}_H^\male \bar{w}_H^\male\}$ is the mean fitness of males during diploid competition
    \end{tablenotes}
  \label{tab:haplotype_growth}
\end{threeparttable}
\\

@@

Table \ref{tab:haplotype_growth} illustrates a number of key points about the invasion of neo-Y and neo-W mutations. 
Firstly, Fisherian sex ratio selection will favour the spread of a neo-Y if the ancestral zygotic sex ratio is biased towards males, $\zeta>1/2$, and vice versa for a neo-W, see terms in square brackets. 
However, the spread of a neo-Y (neo-W) also depends on the male (female) fitness of alleles that they are associated with, see terms in curly brackets. 
Secondly, since a dominant neo-Y is always found in males, the allele frequencies at the neo-Y (\textbf{M}) locus only change in males. 
Therefore, invasion by a neo-Y allele does not involve any female diploid selection terms ($w_{g}^{\female}$). 
Similarly, invasion by a neo-W is driven by the fitness of female gametes and diploids and does not involve any direct selection in male diploids.

>>>>

%\textcolor{red}{I now remove the zygotic sex ratio $\zeta$ from the mean fitnesses. The mean fitnesses will have to be adjusted in a corresponding way.}
%
%\textcolor{red}{CHECK: DO THE HAPLOID MEAN FITNESSES HAVE TO BE IN THE DENOMINATOR? i think so, based on definitions in table S.2 (mmo)}

@@

\begin{threeparttable}[ht]
\centering
\smallskip
%\caption{Parameters determining invasion (equation \ref{eq:charpoly_neoY}) for neo-Y or neo-W alleles}
\caption{Parameters determining invasion of mutant neo-Y and neo-W alleles into an ancestrally XY system}
\begin{tabular}{l}
\hline\hline
   \noalign{\vskip 0.5ex}
   neo-Y ($k=0$) \\ [0.5ex] \hline \noalign{\vskip 1ex}
  $\lambda_{mA}=[1/(2 (1- \zeta))]\{p_{X}^{\female} w_{A}^{\female} w_{A}^{\male} w_{AA}^{\male} + 2(1-p_{X}^{\female}) w_{a}^{\female} w_{A}^{\male} w_{Aa}^{\male} \alpha^\male (1 - R)\}/ \{\bar{w}^\male \bar{w}_H^\female \bar{w}_H^\male \}  $\\ [0.5ex] \noalign{\vskip 0.5ex}
  $\lambda_{ma}=[1/(2(1-\zeta))]\{(1-p_{X}^{\female}) w_{a}^{\female} w_{a}^{\male} w_{aa}^{\male} + 2p_{X}^{\female} w_{A}^{\female} w_{a}^{\male} w_{Aa}^{\male}(1 - \alpha^\male) (1 - R)\}/ \{\bar{w}^\male \bar{w}_H^\female \bar{w}_H^\male \}  $ \\ [0.5ex] \noalign{\vskip 0.5ex}
  $\rho_{mA}=R (1-p_{X}^{\female}) w_{a}^{\female} w_{A}^{\male} w_{Aa}^{\male} \alpha^{\male}/  \{(1-\zeta) \bar{w}^\male \bar{w}_H^\female \bar{w}_H^\male \}   $\\ [0.5ex] \noalign{\vskip 0.5ex}
  $\rho_{ma}=R p_{X}^{\female} w_{A}^{\female} w_{a}^{\male} w_{Aa}^{\male} (1 - \alpha^{\male})/ \{(1-\zeta) \bar{w}^\male \bar{w}_H^\female \bar{w}_H^\male \}  $\\ [1ex] \hline 
  \noalign{\vskip 0.5ex}
  neo-W ($k=1$) \\ [0.5ex] \hline \noalign{\vskip 1ex}
  $\lambda_{mA}=[1/(2 \zeta)]\{\bar{p}^{\male} w_{A}^{\male} w_{A}^{\female} w_{AA}^{\female}+2(1-\bar{p}^{\male}) w_{a}^{\male} w_{A}^{\female} w_{Aa}^{\female}\alpha^\female(1-R)\}/ \{\bar{w}^\female \bar{w}_H^\female \bar{w}_H^\male \} $ \\ [0.5ex] \noalign{\vskip 0.5ex}
  $\lambda_{ma}=[1/(2 \zeta)]\{(1-\bar{p}^{\male}) w_{a}^{\male} w_{a}^{\female} w_{aa}^{\female}+2\bar{p}^{\male} w_{A}^{\male} w_{a}^{\female} w_{Aa}^{\female}(1-\alpha^\female)(1-R)\}/ \{\bar{w}^\female \bar{w}_H^\female \bar{w}\} $ \\ [0.5ex] \noalign{\vskip 0.5ex}
  $\rho_{mA}=R (1-\bar{p}^{\male}) w_{a}^{\male} w_{A}^{\female} w_{Aa}^{\female} \alpha^{\female}/ \{ \zeta \bar{w}^\female \bar{w}_H^\female \bar{w}_H^\male \}  $\\ [0.5ex] \noalign{\vskip 0.5ex}
  $\rho_{ma}=R\bar{p}^{\male} w_{A}^{\male} w_{a}^{\female} w_{Aa}^{\female} (1-\alpha^{\female})/ \{ \zeta \bar{w}^\female \bar{w}_H^\female \bar{w}_H^\male \} $\\ [1ex] 
  \hline \hline 
   \end{tabular}
      \begin{tablenotes}
      \scriptsize
      \item $\bar{p}^{\male}=p_{Y}^{\male}q+p_{X}^{\male}(1-q)$ is the average frequency of the $A$ allele among X- and Y-bearing male gametes.
      \item $R$ is the probability of recombination between loci \textbf{A} and \textbf{M}.
      \item $\zeta$ is the zygotic sex ratio (fraction female)
      \item $\bar{w}^\Hermaphrodite$ is the mean fitness of diploids of sex $\Hermaphrodite$, see Table \ref{tab:meanfitnesses}
      \item $\bar{w}_H^\Hermaphrodite$ is the mean fitness of haploids from sex $\Hermaphrodite$, see Table \ref{tab:meanfitnesses}
%      \item $\bar{w}_H^\female = p_X^\female w_A^\female + (1-p_X^\female) w_a^\female$ is the mean fitness of female gametes during gametic competition
%      \item $\bar{w}_H^\male = \bar{p}^{\male} w_A^\male + (1-\bar{p}^{\male}) w_a^\male$ is the mean fitness of male gametes during gametic competition
%       \item $\bar{w}^\female = \{ p_X^\female w_A^\female (1-q) p_X^\male w_A^\male w_{AA}^\female + (1 - p_X^\female) w_a^\female (1-q) p_X^\male w_A^\male w_{Aa}^\female + p_X^\female w_A^\female (1-q) (1 - p_X^\male) w_a^\male w_{Aa}^\female + (1-p_X^\female) w_a^\female (1-q) (1 - p_X^\male) w_a^\male w_{aa}^\female \} / \{ \bar{w}_H^\female \bar{w}_H^\male\}$ is the mean fitness of females during diploid competition
%       \item $\bar{w}^\male = \{ p_X^\female w_A^\female q p_Y^\male w_A^\male w_{AA}^\male + (1 - p_X^\female) w_a^\female q p_Y^\male w_A^\male w_{Aa}^\male + p_X^\female w_A^\female q (1 - p_Y^\male) w_a^\male w_{Aa}^\male + (1-p_X^\female) w_a^\female q (1 - p_Y^\male) w_a^\male w_{aa}^\male \} / \{ \bar{w}_H^\male \bar{w}_H^\male\}$ is the mean fitness of males during diploid competition
    \end{tablenotes}
  \label{tab:haplotype_growth}
\end{threeparttable}

@@

Table \ref{tab:haplotype_growth} illustrates a number of key points about the invasion of neo-Y and neo-W mutations. 
Firstly, Fisherian sex ratio selection will favour the spread of a neo-Y if the ancestral zygotic sex ratio is biased towards females, $\zeta>1/2$, and vice versa for a neo-W, see terms in square brackets. 
However, the spread of a neo-Y (neo-W) also depends on the male (female) fitness of alleles that they are associated with, see terms in curly brackets. 
Secondly, since a dominant neo-Y is always found in males, the allele frequencies at the neo-Y (\textbf{M}) locus only change in males. 
Therefore, invasion by a neo-Y allele does not involve any female diploid selection terms ($w_{g}^{\female}$). 
Similarly, invasion by a neo-W is driven by the fitness of female gametes and diploids and does not involve any direct selection in male diploids.
